---
title: "ST_558_HW_3"
author: "Stephanie Shuai, Jiayin Liu"
format: html
editor: visual
---

```{r,echo=FALSE}
library(dplyr)
library(tidyverse)
library(readr)
```

# Task 1

**#1. If your working directory is myfolder/homework/, what relative path would you specify to get the filelocated at myfolder/MyData.csv?**

The direct path should be../MyData.csv

**#2.What are the major benefits of using R projects?**

R is good for a lot of task, which includes: 1. Read raw data in or connect to a database. 2. Manipulate data as need. 3.Subset. 4. Create new variables. 5.Summarize data to create meaningful insights 6.Modeling data to make inference or predict outcomes. 7.Communicate our results via dashboards, documents, model files, etc. It is very convinient for data analyst or statistician to do all their task in one single software.

**#3.What is Git and what is GitHub?**

Based on the course material, git is a version control software, while github is an online hosting service for Git-based projects

**#4.What are the two main differences between a tibble and a data.frame?**

First, a tibble prints more nicely. It only shows the first few rows and columns that fit on the screen, while a data.frame may print a lot of issues that may not be necessary

Second, a tibble is stricter and more predictable. For example, it does not automatically change variable types or names as much as a traditional data.frame may.

**#5.Rewrite the following nested function call using baseR’s chaining operator:**

```{r}
iris |>
  as_tibble() |>
  select(starts_with("Petal"), Species) |>
  filter(Petal.Length < 1.55) |>
  arrange(Species)

```

**#6. What is meant by long format data and wide format data? Which do we generally prefer for statistical analysis?**

long format: each row consists of one observation and each column one variable.

wide format data: we may have more than one observation in a given row.

# Task 2- Question about glass data

**#1.Read glass data into R directly from the URL**

```{r}

glass_dt<-read_delim('https://www4.stat.ncsu.edu/online/datasets/glass.data',delim=",",col_names=c('Id','RI','Na','Mg','Al','Si','K','Ca','Ba','Fe','Type_of_glass'))

```

**#2.Overwrite the variable**

```{r}
glass_dt2<-glass_dt %>% mutate(
  Type_of_glass = factor(
    Type_of_glass,
    levels = c(1, 2, 3, 4, 5, 6, 7),
    labels = c(
      "building_windows_float_processed",
      "building_windows_non_float_processed",
      "vehicle_windows_float_processed",
      "vehicle_windows_non_float_processed",
      "containers",
      "tableware",
      "headlamps"
    )
  )
)
```

**#3.Data filer**

```{r}
glass_dt3<-glass_dt2 %>% 
  filter(Fe < 0.2&Type_of_glass %in% c("tableware", "headlamps"))

```

# Task 2- Quetsion about yeast data

**#1.Read glass data into R directly from the URL**

```{r}
yeast_url <- "https://www4.stat.ncsu.edu/~online/datasets/yeast.data"

col_names <- c("seq_name", "mcg", "gvh", "alm", "mit", "erl",
               "pox", "vac", "nuc", "class")

yeast <- read_table( file = yeast_url,col_names = col_names)

# Print out the tibble
yeast
```

**#2. Start a chain that removes the seq_name and nuc columns**

```{r}
yeast_2 <- yeast %>%
  select(-seq_name, -nuc)

# Print out the tibble
yeast_2
```

**#3. Add a column**

```{r}
yeast3 <- yeast_2 %>%
  group_by(class) %>%
  mutate( across(where(is.numeric),mean,.names = "mean_{.col}" ),
          across(where(is.numeric),median,.names='median_{.col}'))

# Print out the tibble
yeast3

```

## Question 3

**#1. Read in the White Wine Excel file**

```{r}
library(readxl)
white_wine <- read_excel("~/Desktop/ST 558/ST 558 HW3/white-wine.xlsx", sheet = 1)
white_wine 
```

**#2. Rename columns using Sheet 2**

```{r}
new_names <- read_excel("~/Desktop/ST 558/ST 558 HW3/white-wine.xlsx", sheet = 2)

colnames(white_wine) <- new_names[[1]]
```

**#3. Add type column for white wine**

```{r}
library(dplyr)

white_wine <- white_wine |> 
  mutate(type = "white")
```

**#4. Read in the Red Wine CSV**

```{r}
library(readr)
red_wine <- read_delim("~/Desktop/ST 558/ST 558 HW3/red-wine.csv", 
                       delim = ";",
                       show_col_types = FALSE)
colnames(red_wine) <- new_names[[1]]

red_wine <- red_wine |> 
              mutate(type = "red")
```

**#5. Combine both datasets**

```{r}
wine <- dplyr::bind_rows(white_wine, red_wine)
```

**#6-9. Combined data object to filter the data**

```{r}
wine |> 
  # Filter high quality and reasonable alcohol 
  filter(quality > 6.5 & alcohol < 132) |>
  # Sort highest quality to lowest
  arrange(desc(quality)) |>
  # Select acid columns, alcohol, type, quality
  dplyr::select(contains("acid"), alcohol, type, quality) |>
  # Add mean and sd of alcohol for each quality level
  group_by(quality) |>
  mutate(
    mean_alcohol = mean(alcohol),
    sd_alcohol = sd(alcohol)
  )
```

## Question 4

**#1. Connect to the database and list all tables**

```{r}

library(DBI)

con <- dbConnect(RSQLite::SQLite(), "~/Desktop/ST 558/ST 558 HW3/lahman.db")

dbListTables(con)
```

**#2. Using the tbl() function and dplyr to return all of the data**

```{r}
library(dbplyr)

tbl(con, "Teams") |>
  filter(yearID == 2015 ) |>
  collect()
```

**#3. Using the sql() function**

```{r}
library(dbplyr)

tbl(con, "Teams") |>
  filter(yearID == 2015 ) |>
  collect()
```

**#4. Inducted players only**

```{r}
hof <- tbl(con, sql("
          SELECT playerID, yearID, category
          From HallOfFame 
          WHERE inducted = 'Y'
          ")) |>
        collect()
hof
```

**#5. Add nameFirst and nameLast from People**

```{r}
hof_names <- tbl(con, sql("
              SELECT h.playerID, h.yearID, h.category, p.nameFirst, p.nameLast
              From HallOfFame h
              JOIN People p ON h.playerID = p.playerID
              WHERE h.inducted = 'Y'
              ")) |>
            collect()
hof_names
```

**#6. Managers win/loss records and win percentage**

```{r}
managers <- tbl(con, "Managers") |>
                select(playerID, G, W, L) |>
                group_by(playerID) |>
                summarize(G_managed = sum(G, na.rm = TRUE),
                  Total_W = sum(W, na.rm = TRUE),
                  Total_L = sum(L, na.rm = TRUE)) |>
                collect() |>
                mutate(win_pct = Total_W/G_managed) |>
                arrange(desc(win_pct))
managers
```

**#7. Join two previous table**

```{r}
hof_names |>
  inner_join(managers, by = "playerID")
```
